### 2018-02-05 thesis/contents/discussions.tex
Updates from ShareLaTeX

The biggest value of tractography to neurosugery is to assit in pre-operative planning. Tractography, in conjuction with 3D anatomical segmentation, can 1) reconstruct white matter pathways that are otherwise invisible, and 2) allow intuitive spatial relationship of key anatomies that are hard to visualise in 2D. We had previously demonstrated these capabilities in the reconstruction of deformed facial/vestibular nerve complex in relation to vestibular schwannomas \cite{Chen2011b,Behan2017}. Tractography and 3D anatomical model ultimately are standard 3D geometries in computer graphics. Therefore advances in 3D printing can be easily adopted in neurosurgery to allow surgeons to physically examine tractography reconstructions. On the otherhand, technologies such as virtual reality (VR) and augmented reality (AR), which are simply different ways to present computer graphics, can allow new ways to empower surgeons in decision making. Tractography can play a key role in bridging computer vision based diagnosis and surgical disease intervention to improve surgical outcomes and improved quality-of-life for patients.

>>>>

The biggest value of tractography to neurosugery is to assist in pre-operative planning. Tractography, in conjuction with 3D anatomical segmentation, can 1) reconstruct white matter pathways that are otherwise invisible, and 2) allow intuitive spatial relationship of key anatomies that are hard to visualise in 2D. We had previously demonstrated these capabilities in the reconstruction of deformed facial/vestibular nerve complex in relation to vestibular schwannomas \cite{Chen2011b,Behan2017}.
Another contribution is to visualize tissue changes from surgical interventions that can otherwise be invisible. There are cases where post-operative GammaKnife ablation target on the trigeminal nerve did not result in dicernable gadonenium enhancement in standard MR, but the nerve FA changes can be seen in the delineated tractography. Therefore tractography can provide both spatial, as well as tissue-based insights to the surgeons to aid in assessment.
Tractography and 3D anatomical model ultimately are standard 3D geometries in computer graphics. Therefore advances in 3D printing can be easily adopted in neurosurgery to allow surgeons to physically examine tractography reconstructions. On the otherhand, technologies such as virtual reality (VR) and augmented reality (AR), which are simply different ways to present computer graphics, can allow new ways to empower surgeons in decision making. Tractography can play a key role in bridging computer vision based diagnosis and surgical disease intervention to improve surgical outcomes and improved quality-of-life for patients.

### 2018-02-05 thesis/contents/tn_sagit.tex
Updates from ShareLaTeX

\subsubsection{Merged Tractography}
All tractography delineations of the same anatomy across all the subjects and groups were merged into a single tractography file for analysis. A software library\footnote{https://github.com/sinkpoint/fascicle} is created by the author to facilitate this process, the steps are as follows: The individual tract files were converted into a SQLite relational database \cite{owens2010sqlite}, and  all points were exported and transformed by applying the subject DWI to T1, and T1 to template transforms using ANTS' applyTransformToPoints command \cite{Avants2009}. The transformed points were then imported back into the  database, and a new template-space tractography model is exported using the transformed point-set, with all its original scalars and subject/group ID associations preserved. Finally, all template-space tractography models across all the subjects are merged into a single file.

@@

The diffusivity measures of CN V reproduced our previous findings \cite{Chen2016a}, which was that cistern, REZ differentiated TN from controls. The consistent regional overlaps in the cistern/REZ region of the right (affected) CN V in MD, RD, AD shows the ability for the GP classifiers to auto-converge to focal white-matter abnormality. The left versus right GP classifier placed left-right differences at the cistern segment, in agreement with the pathophysiology of TN. The regional findings on left versus controls suggest that there is bilateral differences in CN V myelin in unilateral TN, despite the lack of symptoms on the unaffected side. The left regions also situate into the brainstem segments. It is possible that the left CN V is affected by global descending pain modulation, and warrant further investigation.

@@

The TPT delineations showed that decussation pathway tractography is possibly asymmetric. It is not clear if the asymmetry is a neuroanatomically factual, or an artifact from either algorithmic bias or partial volume bias of the underlying DWI. Resolving crossing fiber is still a difficult task in tractography, and while there are validations against MR phantoms, the anatomical validity of these methods can only be extrapolated, and the interpretation must involve anatomical priors from experts.

We included GFA to investigate its potential for measuring deep white-matter diffusivity, as it is the most well-known non-Gaussian diffusivity metric. However S1 results demonstrated that it is not effective in pinpoint deep white-matter abnormalities. GFA performed well in CN V, but not to any substantially degree than the Gaussian-base diffusivity metrics.

>>>>

\subsubsection{Merged Tractography}
All tractography delineations of the same anatomy across all the subjects and groups were merged into a single tractography file for analysis. A software library\footnote{https://github.com/sinkpoint/fascicle} is created by the author to facilitate this process, the steps are as follows: The individual tract files were converted into a SQLite relational database \cite{owens2010sqlite}, and  all points were exported and transformed by applying the subject DWI to T1, and T1 to template transforms using ANTS' applyTransformToPoints command \cite{Avants2009}. The transformed points were then imported back into the  database, and a new template-space tractography model is exported using the transformed point-set, with all its original scalars and subject/group associations preserved. Finally, all template-space tractography models across all the subjects are merged into a single file.

@@

The diffusivity measures of CN V reproduced our previous findings \cite{Chen2016a}, which was that cistern, REZ differentiated TN from controls. The consistent regional overlaps in the cistern/ganglion region of the right (affected) CN V showed the ability for the GP classifiers to auto-converge to focal white-matter abnormality. The left versus right GP classifier placed left-right differences at the cistern/REZ segment, in agreement with the pathophysiology of TN. The regional findings on left versus controls suggest that there is bilateral differences in CN V myelin in unilateral TN, despite the lack of symptoms on the unaffected side. The left regions also situate into the brainstem segments. It is possible that the left CN V is affected by global descending pain modulation, and warrant further investigation.

The TPT delineations showed that decussation pathway tractography is possibly asymmetric. It demonstrated very strong differences from the classifier results, with overlapping differences in the ascending segment. There also seemed to be diffusivity differences in the decussation segments between affected and unaffected sides. The decussaing fibers of the TPT emerges from the TGN, therefore the difference seemed reasonable. It is not clear however, whether the tractography asymmetry of the TPT is neuroanatomically factual, or an artifact from either algorithmic bias or partial volume bias of the underlying DWI. Resolving crossing fiber is still a difficult task in tractography, and while there are validations against MR phantoms, the anatomical validity of these methods can only be extrapolated, and the interpretation must involve anatomical priors from experts.

@@

We included GFA to investigate its potential for measuring deep white-matter diffusivity, as it is the most well-known non-Gaussian diffusivity metric. However results demonstrated that it is not effective in pinpoint deep white-matter abnormalities. GFA did not perform to any substantially degree than the Gaussian-base diffusivity metrics.
